Generating Hubbard Model Solutions from Anderson Impurity Model Solutions

### 2021-08-20 notes.tex
described in more detail how to write down the orthonormal set ket{n}

Due to the translation invariance, the coefficients $C^{0}_{n}$ and $C^{1}_{n}$ are independent of the site indices. The index $n$ actually defines a set of quantum numbers that characterize the state $\ket{n}$. For example it might be a combination of number of particles, parity and total spin angular momentum $(n \equiv n, P, S^z)$. One simple way of choosing the orthogonal basis $\left\{ \ket{n} \right\}$ is to take them from the eigenstates of a system comprising a Hubbard dimer and a conduction bath, decoupled from each other. The set of eigenstates that have the total number of particles equal to $N+1$ will then form the basis in question. Since the dimer is decoupled from the bath, the total wavefunction will be a direct product of a Hubbard dimer wavefunction and a bath wavefunction.

@@

@@

\section{Future goals}
\item Obtain the metal insulator transition of the 2D Hubbard model on the square lattice from our formalism. Compare with what is obtained from DMFT and its improvements. Also, determine the nature of the bath $\Sigma (k,\omega)$ for the transition point.

>>>>

Due to the translation invariance, the coefficients $C^{0}_{n}$ and $C^{1}_{n}$ are independent of the site indices. The index $n$ actually defines a set of quantum numbers that characterize the state $\ket{n}$. For example it might be a combination of number of particles, parity and total spin angular momentum $(n \equiv n, P, S^z)$. One simple way of choosing the orthogonal basis $\left\{ \ket{n} \right\}$ is to take them from the eigenstates of a system comprising a Hubbard dimer and a conduction bath, decoupled from each other. The dimer will of course have 2 sites, which means the conduction bath will have $N-2$ sites, or $N-2$ momentum states. The set of eigenstates that have the total number of particles (combining the dimer and the bath) equal to $N+1$ will then form the basis in question. Since the dimer is decoupled from the bath, the total wavefunction will be a direct product of a Hubbard dimer wavefunction and a bath wavefunction.

@@

\section{Some comments on writing down the orthonormal basis $\left\{ \ket{n} \right\}$}
As mentioned previously, for a parent model of $N$ sites, the orthonormal basis is constructed out of those eigenstates of a decoupled dimer+bath system that have total number of particles equal to $\mathcal{N}_\text{tot} = \mathcal{N}_\text{dim} + \mathcal{N}_\text{bath} = N+1$. The condition on the total number is necessary because that is also the total particile number of the state we are expanding $(c^\dagger_{i\sigma}\ket{\tilde \Phi_0}$. The dimer will have 2 sites, while the rest $N-2$ sites go into the bath.
\\\\
If we are solving the trivial example of $H^H = H^D$ (that is, if we take the Hubbard dimer itself as the full Hubbard model that we are trying to solve), the bath is non-existant. The entire set of $\ket{n}$ for expanding $c^\dagger_{i\sigma}\ket{\tilde \Phi_0}$ is then made up by the $\mathcal{N} = N + 1 = 3, S^z = \frac{1}{2}\sigma$ eigenstates of $H^D$. There are two of those; these two make up the entire orthonormal basis in which we can expand the single-particle excitaions:
\begin{equation}\begin{aligned}
	\left\{ \ket{n} \right\} = \ket{3 \pm \sigma}
\end{aligned}\end{equation}
Similarly, for expanding the hole excitaions $c_{i\sigma}\ket{\tilde \Phi_0}$, we will instead choose the $\mathcal{N} = N - 1 = 1, S^z = -\frac{1}{2}\sigma$ eigenstates of $H^D$. There are again two of those:
\begin{equation}\begin{aligned}
	\left\{ \ket{\overline n} \right\} = \ket{1 \pm -\sigma}
\end{aligned}\end{equation}
\\\\
For a more general $N(>2)$-site model, there will be a bath of $N-2$ momentum states. These momentum states are labelled as $k_1, k_2, ..., k_{N-2}$. Let us designate the set of eigenstates of the Hubbard dimer as \(\left\{ \ket{\Phi^D_i} \right\} \) and those of the bath as \(\left\{ \ket{\Psi^B_{j, \nu^j}} \right\} \). Since the bath is non-interacting, those eigenstates can be written down easily:
\begin{equation}\begin{aligned}
	\left\{ \ket{\Psi^B_{j, \nu^j}} \right\} = \begin{cases} &  j \in [0, 2N-4]\\[-10pt]
	\otimes_{i=1}^j\hat n_{\nu^j_i} = 1:\\[-10pt]
& \nu^j \in \text{all subsets of length $j$ of the set $\left[k_1 \uparrow, k_1 \downarrow, k_2 \uparrow, ..., k_{N-2} \uparrow, k_{N-2} \downarrow\right]$} \end{cases}
\end{aligned}\end{equation}
$j$ defines the number of occupied states in a particular eigenstate of the conduction bath. The string $\nu^j$ runs over all possible set of quantum numbers of a particular length $j$. For example, if we have 3 momentum states in the bath $k_1, k_2, k_3$, then $\nu^2$ would take the values
\begin{equation}\begin{aligned}
	\left\{\nu^2\right\} = \left(k_1 \uparrow, k_1 \downarrow\right),\left(k_1 \uparrow, k_2 \uparrow\right),\left(k_1 \uparrow, k_2 \downarrow\right),\left(k_1 \uparrow, k_3 \uparrow\right),\left(k_1 \uparrow, k_3 \downarrow\right),\left(k_1 \downarrow, k_2 \uparrow\right),\left(k_1 \downarrow, k_2 \downarrow\right),\\
	\left(k_1 \downarrow, k_3 \uparrow\right),\left(k_1 \downarrow, k_3 \downarrow\right),\left(k_2 \uparrow, k_2 \downarrow\right),\left(k_2 \uparrow, k_3 \uparrow\right),\left(k_2 \uparrow, k_3 \downarrow\right),\left(k_2 \downarrow, k_3 \uparrow\right),\left(k_2 \downarrow, k_3 \downarrow\right),\\
\left(k_3 \uparrow, k_3 \downarrow\right)
\end{aligned}\end{equation}
The $j=0$ case means the bath is empty.
The general complete set of orthonormal kets is given by
\begin{equation}\begin{aligned}
	\left\{ \ket{n} \right\} = \sum_{i=1}^{16}\sum_{j=0}^{N-2}\sum_{\nu^j}\ket{\Phi^D_i}\otimes \ket{\Psi^B_{j, \nu^j}}
\end{aligned}\end{equation}
The first twp sums are unconstrained. The index $i$ runs through all 16 eigenstates of the Hubbard dimer, the index $j$ runs through all the possible total number of particles the eigenstate of the bath can have. For each value of $j$, the string $\nu^j$ runs through all possible combinations of length $j$ from the complete set of quantum numbers $\left\{ k\sigma \right\} $.

@@

\section{Future goals}
\begin{enumerate}
\item Obtain the metal insulator transition of the 2D Hubbard model on the square lattice from our formalism. Compare with what is obtained from DMFT and its improvements. Also, determine the nature of the bath $\Sigma (k,\omega)$ for the transition point.
